### 2021-09-06 input/proposal.tex
meow

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issues with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending ($s_{pq}(m) \rightarrow s_{pq}(m') \implies r_q(m) \rightarrow
r_q(m')$). Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ does not deliver $\alpha_d$ and $\delta_d$ since it
already delivered it. The system converges to an inconsistent state
where some \processes assume they belong to a partition while it does
not exist anymore. Naive propagation of $\alpha$ and $\delta$ messages
is insufficient to guarantee consistent partitioning when any \process
can add or delete its partition at any time. \Processes need to
exploit their local knowledge gathered over message receipts.

@@

\begin{proof}
  %% To reach consistent partitioning, the last message delivered by
  %% every process is not stale and comes from its respective best
  %% source: $d(m)$
  Assuming BEF as default propagation behavior where \processes
  deliver and forward the best sum of weights they received. We need
  to prove that despite stale messages, the last delivery of every
  process is the message that followed the shortest path to its
  closest source.
  \begin{asparadesc}
  \item [BEF without Purge:] Figure~\ref{fig:proofA} depicts the
    generic case where a \process $b$ precludes itself and subsequent
    \processes such as $c$ of delivering the message from their
    respective best source $a$. \Process $b$ received and delivered a
    message from a partition that has since then been deleted but with
    a better sum of weights than newly received ones ($\alpha_d^y <
    \alpha_a^{x'}$). It forever discards such a message
    $\alpha_a^{x'}$.
  \item [Purge not followed by BEF:] Even assuming that every process
    immediately removes stale control information $S(\alpha_d^y)
    \implies d_b(\alpha_d^y) \rightarrow d_b(\delta_d)$, but received
    $\alpha^{x'}_a$ earlier $d_b(\alpha^{x'}_a) \rightarrow
    d_b(\alpha^y_d)$, it may never receive it again, hence staying
    with an empty state or another message that did not follow its
    shortest path to its best source.
  \item [Purge then BEF:] Removing \emph{forever} all such stale
    messages about deleted partitions would allow \processes to
    propagate their best up-to-date partition again, eventually
    reaching consistent partitioning together, as stated by
    Theorem~\ref{theo:bef}.
  \end{asparadesc}
\end{proof}

\input{input/figproof.tex}

@@

@@

\begin{algorithm}
  \input{input/algoascast.tex}
  \caption{\label{algo:ascast}\NAME at \Process~$p$ in static networks.}
\end{algorithm}

>>>>

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issues with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending ($s_{pq}(m) \rightarrow s_{pq}(m') \implies r_q(m) \rightarrow
r_q(m')$). Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ does not deliver $\alpha_d$ and $\delta_d$ since it
already delivered it. The system converges to an inconsistent state
where some \processes assume they belong to a partition while it does
not exist anymore. Naive propagation of $\alpha$ and $\delta$ messages
is insufficient to guarantee consistent partitioning when any \process
can add or delete its partition at any time. \Processes need to
exploit their local knowledge gathered over message receipts to allow
up-to-date messages to propagate to relevant \processes.

\begin{proof}
  %% To reach consistent partitioning, the last message delivered by
  %% every process is not stale and comes from its respective best
  %% source: $d(m)$
  Assuming BEF as default propagation behavior where \processes
  deliver and forward the best sum of weights they received. We must
  prove that despite stale messages, the last delivery of every
  process is the message that followed the shortest path to its
  closest source.
  \begin{asparadesc}
  \item [BEF without Purge:] From Figure~\ref{fig:problemD}, a
    \process may receive messages in such an order ($d_b(\alpha_d^1)
    \rightarrow r_b(\delta_a) \rightarrow d_b(\delta_d)$) that it does
    not forward messages required by other \processes
    ($\delta_a$). The last delivery of some nodes ($c$ and $d$) is
    wrong ($\alpha_a^3$ instead of $\delta_a$). \Processes never reach
    consistent partitioning.
    %% Figure~\ref{fig:proofA} depicts the generic case where a \process
    %% $b$ precludes itself and subsequent \processes such as $c$ of
    %% delivering the message from their respective best source
    %% $a$. \Process $b$ received and delivered a message from a
    %% partition that has since then been deleted but with a better sum
    %% of weights than newly received ones ($\alpha_d^y <
    %% \alpha_a^{x'}$). It forever discards such a message
    %% $\alpha_a^{x'}$.
  \item [Purge not followed by BEF:]
    %% also transiting messages
    Assuming that every process removes control information as soon as
    they become stale ($S(\alpha_a^x) \implies d_b(\alpha_a^x)
    \rightarrow d_b(\delta_a)$) and cannot deliver stale messages
    anymore ($S(\alpha_a^x) \implies d_c(m_1) \rightarrow \ldots
    \rightarrow d_c(m_k \not\in \alpha_a)$), staleness does not impair
    up-to-date messages propagation. However, even in such settings,
    depending on receipt order ($d_b(\alpha_a^{x}) \rightarrow
    r_b(\alpha_c^{y : y > x}) \rightarrow d_b(\delta_b)$), a process
    may never receive again, hence deliver and forward a message that
    was received and discarded before ($\alpha_c^{y}$). When such a
    message contains its best up-to-date partition, the \process never
    ends up in its best partition. Therefore, \processes do not reach
    consistent partitioning.
  \item [Purge then BEF:] Assuming the eventual removal of stale
    control information ,

    nothing stops up-to-date control information to propagate, as in
    Theorem~\ref{theo:bef}. Assuming that removal of stale control
    information immediatly triggers a forwarding of best up-to-date
    %% Removing \emph{forever} all such stale messages about
    %% deleted partitions would allow \processes to propagate their best
    %% up-to-date partition again, eventually reaching consistent
    %% partitioning together, as stated by Theorem~\ref{theo:bef}.
  \end{asparadesc}
\end{proof}

%% \begin{figure}
%%   \centering\input{input/figproofA.tex}
%%   \caption{\label{fig:proofA}Stale $\alpha$ messages may
%%       stop other $\alpha$ messages from reaching all \processes ($b$
%%       and $c$) that require it along the shortest path. Consistent
%%       partitioning requires eventual purging of stale messages (see
%%       Figure~\ref{fig:proofB}).}
%% \end{figure}

@@

%% \input{input/figproof.tex}

@@

\begin{figure}
  \centering\input{input/figproofB.tex}
  \caption{\label{fig:proofB}Stale $\alpha$ messages may
      stop $\delta$ messages from reaching processes with the corresponding
      partition. Consistent partitioning requires that all \processes propagate
      $\delta$ messages as often as possible (I'); detect possible inconsistencies
      when parents' $\alpha$ messages contradict history or state (III); notify
      processes of possible inconsistencies (IV)}
\end{figure}

@@

\begin{algorithm}[h]
  \input{input/algoascast.tex}
  \caption{\label{algo:ascast}\NAME at \Process~$p$ in static networks.}
\end{algorithm}
